Annotating, understanding, and predicting long-term video memorability

### 2018-01-29 ARTICLE.tex
Karthik changes 29 january

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{booktabs} % For formal tables
%% perso packages:
\usepackage{epstopdf}
\usepackage{microtype}
\usepackage{tabularx}
\usepackage{multirow}
%\usepackage{arydshln}
\usepackage{subfig}
\usepackage{balance}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{hhline}
\usepackage{multirow}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Annotating, understanding, and predicting long-term video memorability}
\titlenote{Produces the permission block, and copyright information}
\subtitle{Extended Abstract}
\subtitlenote{The full version of the author's guide is available as
\texttt{acmart.pdf} document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AUTHORS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Romain Cohendet}
%\authornote{}
%\orcid{1234-5678-9012}
\affiliation{%
	\institution{Technicolor}
	\streetaddress{}
	\city{Rennes} 
	\state{France} 
	\postcode{}
}
\email{romain.cohendet@technicolor.com}

\author{Karthik Yadati}
\affiliation{%
	\institution{University of Delft}
	\streetaddress{}
	\city{Delft} 
	\state{} 
	\postcode{}
}
\email{N.K.Yadati@tudelft.nl}

\author{Ngoc Khanh Duong}
\affiliation{%
	\institution{Technicolor}
	\city{Rennes} 
	\country{France}}
\email{Quang-Khanh-Ngoc.Duong@technicolor.com}

\author{Claire-Helene Demarty}
\affiliation{%
	\institution{Technicolor}
	\city{Rennes}
	\country{France}
}
\email{Claire-Helene.Demarty@technicolor.com}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Following on from the study of image memorability prediction, which draw continuous attention over the past six years, the computational understanding of video memorability's search field has recently hatched.
The growing number of shared videos foster us to find new ways to make their occurrence the most relevant in our everyday lives.
There is no available dataset of videos annotated in terms of memorability; such dataset may probably result in a launching of the field, as it had been the case for images.
The first goal of the pioneers of this search should be to succeed to propose a protocol to constitute such a dataset. 
%our work:
In this article, we propose a protocol of 700 videos annotated with memorability scores, constitute a dataset, study the quality of the collected data, and computationally modeling memorability to predict it.
%long-term memory

We finally propose a deep-learning model base to predict video memorability, comparing several methods and class of features.

\end{abstract}

@@

%% About the ICMR article on Media retrieval with video memorability
	% focus on "Multimedia content-based retrieval"
	% Length: from 6 to 8 pages + 1 additional page for references (no longer distinction between long and short papers)
	% Submission deadline: 2018/01/20
	% Contributions addressing the challenges of large-scale search and user behavior analysis are especially welcome

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}%nb of sections %2 PAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Opening
Enhancing the relevance of multimedia occurrences in our everyday life requires to imagine new ways to organize -- in particular, to retrieve -- contents.
Like other metrics of video "importance", such as aesthetics or interestingness, memorability can be regarded as a particularly relevant element to help us picking a video among several ones, with the advantage of a possibility to be measured less subjectively.

%from images to video memorability
The computational understanding of video memorability's search field follows on from the study of image memorability prediction which has attracted increasing attention from the seminal work of Isola \textit{et al.} released in 2001 \cite{isola_2011_makes}.
Recently, models of image memorability prediction achieved very good results with the introduction of deep learning to address the challenge of image memorability prediction \cite{khosla_2015_understanding,baveye_2016_deep,squalli_2017_deep}.
This success resulted on the extension of the challenge to videos.
Video memorability prediction is however a new field, and there are only two studies, to our knowledge, that addressed this issue \cite{han_2015_learning,shekhar_2017_show}.

%No dataset available
At least two important problems could explained this scarcity of studies.

%1st point: no available dataset
Firstly, there is no available dataset to train and test models.
The fact that there exist no dataset to modeling video memorability is the most serious obstacle which prevent video memorability prediction to take off.
This should be the very first goal of pioneers of this new research field: to provide data to community.
For images, the pioneers created and made downloadable such a database \cite{isola_2011_makes}, which enable to made image memorability research flourishing.
The second released database, far larger, launched new possibility for deep learning and enable to obtained the aforementioned really good results \cite{khosla_2015_understanding}.
We could expect a similar success for video memorability prediction research.
%A dataset as first objective to launch VM prediction --> Goal of this paper is to participant in the establishment of an adapted protocol
We think that the existence of such a database with high quality data could have the importance for video memorability that the ones of \cite{isola_2011_makes} and \cite{khosla_2015_understanding} had on image memorability.%in particular for deep learning algortihm which perform the best for images

%second point: difficulty to define video unity and VM
This drives us to the second point: the difficulty to build such a dataset.
The principle difficulty to define what video memorability and unity should be, and a protocol to measure items according to these definition.

If we consider the studies on image memorability prediction, we could see that, with their database, the cardinal points defined in the very first study \cite{isola_2011_makes} were widely accepted by the community without re-interrogation of what memorability -- of what the models -- really predicted.
First, it is because Isola \textit{et al.} find a proper and objective way to measure memory performance at a large scale; but it remains that how the authors proceeded is just one way to measure memorability.

From what we can imagine that such a dataset for VM will probably have an important influence on future research, one should define very carefully what memorability of video is and what the video as unity means.
It requires to define the cardinal points that will guide -- and constraint -- researchers that will used the dataset.

First, if image in image memorability was obvious, video in video memorability is less obvious.
Videos are more complex than images. Contrary to images, videos do not constitute clearly defined units, but have supplementary dimensions -- sound and movement -- that makes difficult the definition of what a video is.
%from ahn et al. Videos contain a synthesis of visual, aural, textual, and 	motion information and generally convey relatively richer 	media content than images.
By an extreme example, is it possible for a model to transfer his capacity to predict memorability from movies-like videos to Youtube-like videos?

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Previous work} %Critics of the 2 previous studies
%Critics of studies on image memorability : the problem of sustainable long-term memorability
The work on image memorability prediction, initiated by \cite{isola_2011_makes}, produced great results \cite{khosla_2015_understanding,baveye_2016_deep,squalli_2017_deep}.
The availability of two large datasets \cite{isola_2011_makes,khosla_2015_understanding} has been critical in this achievement. 
However, possibly constrained by the difficulty in conducting long crowdsourcing studies, the authors measured memory performance a few minutes after memorization to obtain their memorability annotations.
As shown in \cite{cohendet_2016_prediction}, this could be a problem if we conceive that memorability reflects a lasting memory performance.
Indeed, it has long been shown that long-term memories continue to change long after their memorization through an ongoing process called consolidation \cite{mcgaugh_2000_memory}.
In particular, the early work of the psychologist Ebbinghaus showed that the drop in long-term memory performance in recall is particularly strong in the few days immediately following the memorization, after which it asymptotically tends its the almost permanent long-term performance \cite{ebbinghaus_1913_memory}.
Because several factors are susceptible to influence the consolidation process (such as the emotional content, the semantics, the attention...), which is not a linear decrease for all the memories, it follows that the order of memorability of images is susceptible to change from some minutes to a day after memorization, as it has been found in \cite{cohendet_2016_prediction} with emotional image from of the International Affective Picture System \cite{lang_2005_international}.
A protocol to collect memorability annotations for videos would benefit from the capacity to measure what we will later refer to as "long-term memorability".

%Critics of Han et al.
To our knowledge, the first attempt at predicting video memorability had been proposed by \cite{han_2015_learning}. 
The authors partially adapted the protocol proposed by \cite{isola_2011_makes} to measure image memorability for videos.
In contrast to the "memory game" proposed by Isola \textit{et al.} to collect memorability data, their protocol is however much heavier. 
They used a classical recognition task to measure memory for videos, which consists in two steps: a free viewing task, followed two days later by a recall task.
We can infer from the sparse information authors gave that the task duration, for each of the 20 participants, was about 24 hours (by taking an average video duration half-way between 15 and 30 sec), spread over 10 sessions (5 free viewing and 5 recall tasks) of about 2 hours each.
Authors used the same proportion of fillers (i.e. non repeated videos) in the free viewing and recall tasks (i.e. 4/5 of fillers and 1/5 of repeated videos named targets) to, according to them, guarantee that viewers were unaware of targets.
If it was mandatory in \cite{isola_2011_makes} for which encoding and recall tasks were interlaced, there is here a way to alleviate the task without impacting its quality; indeed, reducing the number of fillers in free viewing task would have very little impact on memorability scores (often authors even use only material interrogated later in learning/free viewing task).
Furthermore, as pointed by \cite{shekhar_2017_show}, the long time span of the experiment makes difficult the generalization of this protocol, in particular to build an extensive dataset. Regarding the prediction part, we also agree with these authors to say that the method would be difficult to generalize as it used fMRI measurements for predicting memorability.

%Critics of Shekhar et al.
Another earlier approach was the one of \cite{shekhar_2017_show}.
The participants performed a crowdsourcing experiment consisted of a free viewing task where they saw a sequences a videos, followed by a recall task for which they had to answer textual question (such as: "Did you see a man juggling?", "Did you see a car on road?").
The major drawback of the study comes from the use of questions instead of a recognition task.
Indeed, the memorability scores computed for the videos may reflect not only the differences in memory performances but also the differences between the questions in terms of difficulty, ease of understanding (in particular for the non-anglophone people that works with Amazon Mechanical Turk), of imaging, ease to retrieve a scene by some words more the by others... Especially since the authors took into account the response time of the participants to calculate the memorability scores of the videos. One will note that this potential bias could have also affect the measure of inter-human consistency.
Furthermore, the questions are handcrafted, and the choices for types of questions and videos are very limiting (e.g. the question "Did you see a car on road?" implies that in a whole experimental session you can have just one car on a road), and the authors manually ensured that no textual questions nor videos in a sequence were similar in content. This makes difficult the generalization of this protocol to build an extensive dataset.
%%% ATT!! The have tested complexity of textual questions, but there is the other problems! mesure insufisante pour rendre compte de la difficulté réelle des questions.

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Goals of our study}
%Conclusion critics
It appears from the review of previous work that video memorability prediction research is in its early stage.
We still need to find a proper way to collect quality data in order to build a large dataset to bring video memorability prediction research to a new level.
%1/ we want a protocol
The present study aims to be part of these efforts to reach such a way.
As aforementioned, we also want this protocol to measure a "long-term" memorability.
This can be achieved with a study that have a long time span, where learning and recall tasks are seperated by a certain amount of time, or by measuring a memory created prior to the experiment.
Because crowdsourcing is a priori the only practicable way to experimentally collect great amount of annotations, we focused on the last choice more suitable for crowdsourcing as it requires only one connexion to the measurement application.
%2/ we want an all-around paper
To assess the quality of the data produced by the protocol, we analyzed the data obtained from different perspectives, in particular in order to shed light on some potential human or task-related factors susceptible to influence memorability as it is measured by the protocol, that are numerous \ref{cohendet_2016_using}.
%3/ we want to meet the challenge of video memorability prediction
Thanks to the data collected, we participate in the VM prediction challenge.
%Conclusion intro
The structure of the rest of the article follows these points: part 2 concerns the ground truth collection, part 3 the study of the collected data and the calculation of the memorability scores of the sequences, part 4 predicting memorability, and a conclusive part 5 where we talked about intended future work related to video memorability prediction.

%%%%%%% POTENTIALLY %%%%%%%
% Response time: must we exmploit it to calculate memorability scores?
 %Because of \cite{shekhar_2017_show} and other papers in images which suppose that short time response equal to more memorable ALORS qu'ils n'ont pas pour tâche de répondre le plus rapidement possible [vérifier].
 %La répartition du temps de réponse pourrait ne pas linéraire.
	%We hypothesized that the most memorable the sequence, the faster the participant will answer. Parce que les video les mieux représentés dans l'esprit sont les plus accessibles et que les participants ne veulent pas louper la préponse quand ils sont sûrs. % Citer papiers - s'il en existe -- où les auteurs exploitent le temps de réponse pour corriger les scores de mémorabilité;

% Personalization of the data
 %Taking into account human factors %see results Logistic regression vs. SVM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Construction of the dataset}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Our first step was to implement a protocol to measure the long-term memorability of 700 video sequences, extracted from 100 mainstream movies.
The main characteristic of the protocol, in contrast with previously proposed methods to collect image and video memorability annotations, is the absence of learning/free viewing task, replaced by a questionnaire designed to collect information about the participants' prior memory.
The way of proceeding aimed at measuring memory performance after a significant retention period.

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Dataset}
%% films and sequences
We mixed popularity and genres to establish our list of movies.% occidental movies, except for Slumdog millionaire)%to check%(list added to the appendix [ref_appendix])
Then, we manually selected seven sequences of 10 seconds from each movie (for a total of 700 sequences). %with HD or dvdrip quality, in orginal (i.e.english) version, without without subtitles [add if results HD vs. DVDrip]

%cuts
We made efforts to maintain a high intra-sequence semantic cohesion; in particular, we did not cut through meaningful vectors nor we agglomerated shots that belonged to different scenes.
Indeed, we think the images' characteristic by which they constitute obvious semantic units is an important element in the capacity of the prediction models to generalize.
By contrast, the meaning a sequence conveyed was susceptible to change depending on where we decided to cut the movie.

Because the semantics is linked to the memorability of images \cite{isola_2014_makes}, so probably to videos, therefore their memorabilities, we chose this definition susceptible to be extensive to predict other dataset.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{booktabs} % For formal tables
%% perso packages:
\usepackage{epstopdf}
\usepackage{microtype}
\usepackage{tabularx}
\usepackage{multirow}
%\usepackage{arydshln}
\usepackage{subfig}
\usepackage{balance}
\usepackage{caption}
\usepackage{graphicx}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Annotating, understanding, and predicting long-term video memorability}
%\titlenote{}
%\subtitle{}
%\subtitlenote{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AUTHORS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%author 1
%\author{Romain Cohendet}
%\authornote{}
%\orcid{1234-5678-9012}
%\affiliation{%
%	\institution{Technicolor}
%	\streetaddress{Technicolor Research \& Licensing, 975 Avenue des Champs Blancs, 35576 Cesson-Sevigne, France}
%	\city{Rennes} 
%	\state{France} 
%	\postcode{}
%}
%\email{romain.cohendet@technicolor.com}

%author 2
%\author{Karthik Yadati}
%\affiliation{%
%	\institution{University of Delft}
%	\streetaddress{}
%	\city{Delft} 
%	\state{} 
%	\postcode{}
%}
%\email{N.K.Yadati@tudelft.nl}

%author 3
%\author{Ngoc Khanh Duong}
%\affiliation{%
%	\institution{Technicolor}
%	\city{Rennes} 
%	\country{France}}
%\email{Quang-Khanh-Ngoc.Duong@technicolor.com}

%author 4
%\author{Claire-Helene Demarty}
%\affiliation{%
%	\institution{Technicolor}
%	\city{Rennes}
%	\country{France}
%}
%\email{Claire-Helene.Demarty@technicolor.com}

%%%%%%%%%%%%
%%% for blind review %%%
%author 1
\author{Removed for review}
%author 2
\author{Removed for review}
%author 3
\author{Removed for review}
%author 4
\author{Removed for review}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
The growing number of video contents on the Internet encourages us to find new ways to make more relevant their occurrences in our everyday life.
In this context, memorability can be regarded as a useful metric of video importance to help us make a choice between competing videos.
However, the research on computational understanding of video memorability is in its early stages.
There is no available dataset for modelling purposes, and the few previous attempts provided protocols not generalizable to collect data at a large scale.
Furthermore, the video computational features valuable to build a robust memorability predictor remain largely undiscovered.
In this article, we propose a new protocol to collect long-term memorability annotations, that we use to measure memory performances of 104 participants from weeks to years after memorization.
We then analyze the data collected for 660 videos by focusing on its quality, and test it against different characteristics such as response time, duration of memory retention and repetition of visualization.
We finally conduct an extensive feature analysis, comparing several methods and classes of features, to propose a computational model for the prediction of video memorability.
\end{abstract}

@@

%% About the ICMR article on Media retrieval with VM
	% focus on "Multimedia content-based retrieval"
	% Length: from 6 to 8 pages + 1 additional page for references (no longer distinction between long and short papers)
	% Submission deadline: 2018/01/20
	% Contributions addressing the challenges of large-scale search and user behavior analysis are especially welcome

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}%nb of sections %2 PAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%opening
Enhancing the relevance of multimedia occurrences in our everyday life requires to imagine new ways to organize -- in particular, to retrieve -- digital contents.
Like other metrics of video "importance", such as aesthetics or interestingness, memorability can be regarded as useful to help us make a choice between competing videos.
In addition, memorability has the advantage of being clearly definable and objectively measurable (i.e., using a measure that is not influenced by the observer's personal judgment).
Thus, memorability has initially be defined as the probability for an image to be recognized \textbf{a few minutes} after a single view, when presented amidst a stream
of images \cite{isola_2011_makes}.
This definition has been widely accepted within subsequent work (e.g., \cite{mancas_2013_memorability,kim_2013_relative,celikkale_2013_visual,khosla_2015_understanding,lahrache_2016_bag}).

%from images to VM
The computational understanding of video memorability (VM) follows on from the study of image memorability prediction which has attracted increasing attention from the 2001 seminal work of Isola \textit{et al.} \cite{isola_2011_makes}.
With the recent introduction of deep learning to address the challenge of image memorability prediction, models also achieved very good results \cite{khosla_2015_understanding,baveye_2016_deep,squalli_2017_deep}.
As a result of this success, researchers have recently extended this challenge to the videos.
However, to the best of our knowledge, only two available studies focused on VM prediction \cite{han_2015_learning,shekhar_2017_show}.

%explain the scarcity of studies on VM
Several problems could explained this scarcity of studies.
Firstly, there is no publicly available dataset to train and test models.
This is probably the most serious obstacle to the rapid expansion of the VM prediction's search field.
Accordingly, to provide researchers with ground truth data should be our very first objective, similarly with the work of \cite{isola_2011_makes} and \cite{khosla_2015_understanding} which have enabled research on image memorability to flourish.
The second point is closely related to the first one: there is no widely accepted definition of VM.
The previous attempts to predict VM \cite{han_2015_learning,shekhar_2017_show} were based on different measures of memorability.
Furthermore, contrary to images, videos do not constitute clearly defined units.
They have supplementary dimensions -- sound and movement -- that critically contribute to the semantic and emotional information conveyed.
If harmonized, the videos used and the way memorability is measured will have a critical impact on what we will understand by VM.
The definition of image memorability by \cite{isola_2011_makes} had a great impact on subsequent work as aforementioned.
But it also inevitably limited researchers.
In particular, in previous research image memorability corresponds to memory performances measured only a few minutes after memorization.
But these might be poor predictors of longer term memory performances (at least in some instances; e.g., for emotional images \cite{cohendet_2016_prediction}).
Thus, VM data would benefit from a protocol that would measure lasting long-term memory performance.

%modelling
Regarding modelling, our capacity to propose efficient computational models is also important to meet the challenge of VM prediction. 
The previous attempts at predicting VM \cite{han_2015_learning,shekhar_2017_show} shed light on several features which have a predictive power of VM.
However, videos contain a great volume of data which can be used for feature extraction purposes, and the work is far for complete.

%objectives of the present study
The aim of this article is to participate in the expansion of the VM prediction emerging search field.
In particular, we:
\begin{itemize}
	\item propose a new protocol that measures very long-term memory performances (from weeks to years after memorization) to collect quality ground truth data (section 2);
	\item assess and analyze the collected data to better understand VM (section 3);
	\item model VM in an attempt to understand which computation features are better for this task (section 4).
\end{itemize}
In the next subsection, we review previous work on image and video memorability prediction, focusing on protocol and modelling questions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Previous work}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Measurement of memorability}
%Images work -- towards a long-term memorability
Almost all studies on image memorability prediction made use of one or the other of the two available large datasets designed specifically to meet this challenge \cite{isola_2011_makes,khosla_2015_understanding}.
To build these datasets, the authors, possibly constrained by the difficulty in conducting long crowdsourcing studies, measured memory performance a few minutes after memorization to obtain their memorability annotations.
As proposed in \cite{cohendet_2016_prediction}, this could be a problem if we conceive that memorability reflects a lasting memory performance.
Indeed, it has long been shown that long-term memories continue to change long after their memorization through an ongoing process called consolidation, which lasts weeks to years \cite{mcgaugh_2000_memory}.
%In particular, the early work of the psychologist Ebbinghaus showed that the drop in long-term memory performance in recall follow an exponential decay, and is particularly strong in the few days immediately following the memorization \cite{ebbinghaus_1913_memory}.
Because several factors influence the consolidation process (e.g., emotions, sleep, re-evocation), which does not equally affect all memories, the order of memorability ratings measured for videos is susceptible to change over time \cite{cohendet_2016_prediction}.
A protocol to collect memorability annotations would benefit from the capacity to capture long-lasting memory performances, averaged to obtain what we will later refer to as "long-term memorability".

%Han et al.
To our knowledge, the first attempt at predicting VM had been proposed by \cite{han_2015_learning}. 
The authors partially adapted the protocol proposed by \cite{isola_2011_makes} to measure image memorability for videos.
In contrast to the "memory game" proposed by Isola \textit{et al.} to collect memorability data, their protocol is however much heavier.
They used a classical recognition task to measure memory for videos, which consists in two steps: a free viewing task, followed two days later by a recall task.
The task duration, for each of the 20 participants, was about 24 hours, spread over 10 sessions (five free viewing tasks and five recall tasks) of about two hours each.
The authors used the same proportion of fillers (i.e., non repeated videos) in the free viewing and recall tasks (i.e., 4/5 of fillers and 1/5 of repeated videos named targets) to, according to them, guarantee that viewers were unaware of targets.
If it was mandatory in \cite{isola_2011_makes} for which encoding and recall tasks were interlaced, there is here a way to alleviate the task without impacting its quality; indeed, reducing the number of fillers in the free viewing task would have very little impact on the memorability scores (often authors even use only material interrogated later in the learning/free viewing task).
Furthermore, as pointed by \cite{shekhar_2017_show}, the long time span of the experiment makes difficult the generalization of this protocol, in particular to build an extensive dataset.
Furthermore, authors measure memory after two days, but the consolidation process lasts weeks to years: it would be interesting to collect even longer term memory performance measures.

%Shekhar et al.
Another earlier approach was the one of \cite{shekhar_2017_show}.
The participants performed a crowdsourcing experiment consisted of a free viewing task where they saw a sequence a videos, followed by a recall task for which they had to answer textual question (such as: "Did you see a man juggling?", "Did you see a car on a road?").
The major drawback of the study comes from the use of questions instead of a classic visual recognition task.
Indeed, the memorability scores computed for the videos may reflect not only the differences in memory performances but also the differences between the questions in terms of difficulty
The authors have tested complexity of textual questions, but their measure is not sufficient to represent the real difficulty of the question in its complexity (ease of understanding, in particular for the non-anglophone people that works with Amazon Mechanical Turk, ease of imaging, ease to retrieve a scene by some words more the by others...).
Especially since the authors took into account the response time of the participants to calculate the memorability scores of the videos.
One will note that this potential bias could have also affected the measure of inter-human consistency.
Furthermore, the questions are handcrafted, and the choices for types of questions and videos are very limited (e.g., the question "Did you see a car on a road?" implies that in a whole experimental session one can have just one car on a road).
The authors also manually ensured that no textual questions nor videos in a session were similar in content.
This makes it difficult to generalize this protocol to the construction of an extensive dataset.

\subsection{Memorability modelling}
%intro on focus on features to discover
Previous attempts at predicting image and video memorability highlighted quite a few features that correlates with memorability.

%image memorability prediction
The pioneering work of Isola \textit{et al.} focused primarily on building computational models to predict image memorability from low-level visual features \cite{isola_2011_makes}.
It appeared from this first attempt that it is possible to predict to some extent the degree of picture's memorability.
Several characteristics have been found to be relevant for predicting memorability in subsequent work, for example saliency features \cite{mancas_2013_memorability}, interestingness and aesthetic \cite{isola_2014_makes}, or emotions \cite{khosla_2015_understanding}.
The best results was finally obtained by using fine-tuned deep features, which outperformed all other features in \cite{khosla_2015_understanding}, reaching a rank correlation of $.64$ which is near human consistency ($.68$) measured for ground truth collected in the study.
This result was confirmed in \cite{baveye_2016_deep,squalli_2017_deep}.

%VM prediction
Regarding the VM prediction, Han \textit{et al.} propose a method which combines the power of audiovisual and fMRI-derived features \cite{shekhar_2017_show}.
They preliminary built a computational model learned with fMRI features, which supposedly convey the brain activity of memorizing videos, which enable them to finally predict VM without fMRI scans.
However, the method would be difficult to generalize

Shekhar \textit{et al.} conducted a performance analysis of several computationally extracted features before building their memorability predictor \cite{shekhar_2017_show}.
The analysis encompassed C3D deep learning feature for video classification, video semantics obtained thanks to video captioning method, saliency features, dense trajectories, and color features.
They found that the most predictive feature combination used video semantics, spatio-temporal, saliency and color features.
The feature that performed the best when tested alone was video semantics.
The authors used a captioning method to generate a semantic description of the videos.
The text generated by this method was then processed by a recursive auto-encoder network that outputted a 100-dimensional representation of the videos.
Due to the particularity of the dataset of Shekhar \textit{et al.} -- that is, their aforementioned use of questions to measure memorability --, it would be interesting to confirm if this combination works equally well on another dataset, and in particular if images captioning features also perform the best.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Memorability dataset construction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Video collection}
%why did we choose movie
We wanted our protocol to measure memory performance after a significant retention period.
This can be achieved either with a longitudinal study, or by measuring a memory created prior to the experiment.
We chose the latter because it enabled us to immediately measure very long-term memory.
Thus, the main characteristic of the proposed protocol, in contrast with previous work, is the absence of learning (often, free viewing) task, replaced by a questionnaire designed to collect information about the participants' prior memory.
